Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-19 thesis/chapters/ch_introduction.tex
reformatting some of the parenthetical citations

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, all of the elements in the Universe are produced in nuclear reactions associated with stellar evolution, either in the cores of stars, stellar winds, asymptotic giant branch (AGB) winds of low mass stars, and more exotic sites like neutron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012,AndrewsMartini2013}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{Searle1971,Shaver1983,Belfiore2017,SanchezMenguiano2017}, and detailed stellar abundances in nearby dwarf galaxies with spectroscopically resolved stars (see the review in \cite{Tolstoy2009}). Yet in spite of increased number and quality of observations of galactic chemical evolution, developing a complete model of galactic chemical evolution is still a daunting task.

@@

Historically, codes fall into one of two camps: 1) Eulerian grid-based codes, including the popular codes / algorithms commonly in use (in some way) today such as \code{Zeus} \citep{StoneNorman1992}, \code{FLASH} \citep{FLASH}, \code{RAMSES} \citep{Teyssier2002}, \code{Athena} \citep{Athena}, \code{ART-II} \citep{Rudd2008}, and \code{Enzo} \citep{Enzo2014}, and 2) particle-based Lagrangian methods known as smooth particle hydrodynamics (SPH), such as \code{Gadget} \citep{Springel2005}, \code{PKDGRAV-2} \citep{Stadel2001}, \code{Gasoline} \citep{Wadsley2004}, and \code{Changa} \citep{Menon2015}. However, recent codes blur the lines between these distinctions with new algorithms for solving the fluid equations on a moving-mesh (e.g. \code{AREPO} \citep{Springel2010}), or meshless finite-mass / finite-volume methods (such as those implemented in \code{Gizmo} \citep{Hopkins2015}). Traditionally these codes were designed to run exclusively on CPUs, but significant work has been made recently to offload various portions onto GPUs (for example \code{GAMER} and \code{GAMER-2} \citep{Schive2010,Schive2018}, \code{GPUGAS} \code{Kulikov2014}, and \code{CHOLLA} \citep{Schneider2015}). Given the large variance in numerical methods, recent studies have examined the differences between these numerical implementations \citep[e.g.][]{Agertz2007}, including large-scale code comparison projects \citep[e.g.][]{AGORA,AGORA2}. While developing even a notion of which code is ``correct" is an ill-posed problem, these studies have allowed for improvement across implementations and an insight into how numerical methods themselves drive uncertainty in our understanding of astrophysical problems.

@@

The Milky Way and its local environment is host to a diverse collection of dwarf galaxies. The properties of these galaxies have been nicely summarized in a recent review \citep{Tolstoy2009}, and again in \cite{McConnachie2012}. However, the number of known Local Group dwarf galaxies has increased dramatically since that time, particularly for the Milky Way, whose known satellite population has more than doubled. This is due in large part to tremendous improvements in our ability to detect galaxies on the faintest end of the luminosity function, called ultrafaint dwarf galaxies (UFDs, \cite{Willman2005}).\footnote{See \cite{Simon2019} for a recent review}. The Dark Energy Survey \citep[e.g.][]{Drlica-Wagner2015} and Hyper Supbrime-Cam \citep[e.g.][]{Greco2018} have made tremendous advances to this end, with the upcoming Large Synoptic Survey Telescope expected to produce an even greater increase in the known population of faint galaxies in the Local Group \citep{Haynes2019,Weisz2019}.

@@

We dedicate this section to a specific Local Group dwarf galaxy, Leo P, as its observed properties motivate the initial conditions of the simulations developed in this \dissertation. This galaxy is remarkable for being one of the lowest mass dwarf galaxies with observed, ongoing star formation. It is located at a distance of $1.62\pm 0.15$~Mpc \citep{McQuinn2015a}, and was first characterized as part of the ALFALFA survey \citep{Giovanelli2013,Rhode2013,Skillman2013,McQuinn2013,Bernstein-Cooper2014}. These studies have characterized its star formation history as low and roughly continuous over the age of the Universe ($<$SFR$>$ = $4.3\times 10^{-5}$~M$_{\odot}$~yr$^{-1}$, \cite{McQuinn2015a}), its stellar mass ($M_{*} = 5.7 \times 10^{5}$~M$_{\odot}$, \cite{McQuinn2013}), its \hi and dynamical mass ($M_{\rm HI}(r<r_{\rm HI}) = 9.5\times 10^{5}$~M$_{\odot}$, $M_{\rm dyn}(r<r_{\rm HI}) = 2.6 \times 10^{7}$~M$_{\odot}$, $r_{\rm HI} \sim 500$~pc, \cite{Bernstein-Cooper2014}), and gas-phase abundances from an \hii region (12 + log(O/H) = 7.17$\pm$0.04 \cite{Skillman2013}). HI, optical, and H$\alpha$ observations of Leo P are given in Figure~\ref{intro:fig:Leo P}.

@@

One of the most common ways to model and understand the stellar populations of galaxies in high-dimensional chemical space is through simplified one-zone (or many-zone) models. This treatment extends back many decades \citep[e.g.][]{Schmidt1963,TalbotArnett1971,Lynden-Bell1975}, where galaxies and their total metal abundances were treated in closed-box systems of gas and stars. Over the intervening decades these models have improved, including prescriptions for gas outflow (a ``leaky-box") or gas inflow (an ``accreting box"), or both (a ``bathtub" model, e.g. \cite{FinlatorDave2008,Bouche2010}). The assumptions that go into these models vary dramatically with complexity. Simple models can assume, for example, that long-lived low mass stars do not contribute to metal enrichment while metals from massive stars are instantaneously recycled into the galaxy's gas reservoir (i.e. ignoring stellar lifetimes) or that all metals mix instantly and homogeneously. More complex models account for individual stellar lifetimes (to some degree), or may try and account for some aspects of inhomogeneous mixing by constructing multi-zone models that may separate a galaxy in radial bins (with or without mixing between bins), may account for a hot and cold component separately (with mixing between the two), or account for separate ISM and CGM components. The output of these models is typically the mean metal abundance evolution within each of the tracked components. There is generally no self-consistent prescription to account for the scatter in any abundance relationship in these models, which, when needed, is often added ad-hoc in post-processing. Regardless of the assumptions that go into these models, they are powerful tools that can be used to probe the general physical processes that govern galactic chemical evolution, and explore the vast, uncertain parameter space associated with the models that characterize these processes \citep[e.g.][]{Cote2017a}.

>>>>

A substantial slice of modern astrophysics has been devoted towards understanding the chemical evolution -- the abundances of individual elements in space and time -- of the Universe. With the exception of H, He, and trace amounts of light elements, all of the elements in the Universe are produced in nuclear reactions associated with stellar evolution, either in the cores of stars, stellar winds, asymptotic giant branch (AGB) winds of low mass stars, and more exotic sites like neutron star - neutron star mergers (NS-NS). See \cite{Nomoto2013}, \cite{Thielmann2017}, and \cite{Frebel2018} for recent reviews on this topic as it applies to studies of galactic and stellar evolution. Much work has been done in studying the global metallicity evolution of galaxies through the mass-metallicity relationship \citep[e.g.][]{Lequeux1979,Tremonti2004,Lee2006,Zahid2012,AndrewsMartini2013}, the metallicity gradient in our own and nearby galaxies \citep[e.g.][]{Searle1971,Shaver1983,Belfiore2017,SanchezMenguiano2017}, and detailed stellar abundances in nearby dwarf galaxies with spectroscopically resolved stars \citep[see the review in][]{Tolstoy2009}. Yet in spite of increased number and quality of observations of galactic chemical evolution, developing a complete model of galactic chemical evolution is still a daunting task.

@@

Historically, codes fall into one of two camps: 1) Eulerian grid-based codes, including the popular codes / algorithms commonly in use (in some way) today such as \code{Zeus} \citep{StoneNorman1992}, \code{FLASH} \citep{FLASH}, \code{RAMSES} \citep{Teyssier2002}, \code{Athena} \citep{Athena}, \code{ART-II} \citep{Rudd2008}, and \code{Enzo} \citep{Enzo2014}, and 2) particle-based Lagrangian methods known as smooth particle hydrodynamics (SPH), such as \code{Gadget} \citep{Springel2005}, \code{PKDGRAV-2} \citep{Stadel2001}, \code{Gasoline} \citep{Wadsley2004}, and \code{Changa} \citep{Menon2015}. However, recent codes blur the lines between these distinctions with new algorithms for solving the fluid equations on a moving-mesh \citep[e.g. \code{AREPO}][]{Springel2010}, or meshless finite-mass / finite-volume methods \citep[such as those implemented in \code{Gizmo}][]{Hopkins2015}. Traditionally these codes were designed to run exclusively on CPUs, but significant work has been made recently to offload various portions onto GPUs, such as \code{GAMER} and \code{GAMER-2} \citep{Schive2010,Schive2018}, \code{GPUGAS} \citep{Kulikov2014}, and \code{CHOLLA} \citep{Schneider2015}. Given the large variance in numerical methods, recent studies have examined the differences between these numerical implementations \citep[e.g.][]{Agertz2007}, including large-scale code comparison projects \citep[e.g.][]{AGORA,AGORA2}. While developing even a notion of which code is ``correct" is an ill-posed problem, these studies have allowed for improvement across implementations and an insight into how numerical methods themselves drive uncertainty in our understanding of astrophysical problems.

@@

The Milky Way and its local environment is host to a diverse collection of dwarf galaxies. The properties of these galaxies have been nicely summarized in a recent review \citep{Tolstoy2009}, and again in \cite{McConnachie2012}. However, the number of known Local Group dwarf galaxies has increased dramatically since that time, particularly for the Milky Way, whose known satellite population has more than doubled. This is due in large part to tremendous improvements in our ability to detect galaxies on the faintest end of the luminosity function, called ultrafaint dwarf galaxies \citep[UFDs][]{Willman2005}.\footnote{See \cite{Simon2019} for a recent review}. The Dark Energy Survey \citep[e.g.][]{Drlica-Wagner2015} and Hyper Supbrime-Cam \citep[e.g.][]{Greco2018} have made tremendous advances to this end, with the upcoming Large Synoptic Survey Telescope expected to produce an even greater increase in the known population of faint galaxies in the Local Group \citep{Haynes2019,Weisz2019}.

@@

We dedicate this section to a specific Local Group dwarf galaxy, Leo P, as its observed properties motivate the initial conditions of the simulations developed in this \dissertation. This galaxy is remarkable for being one of the lowest mass dwarf galaxies with observed, ongoing star formation. It is located at a distance of $1.62\pm 0.15$~Mpc \citep{McQuinn2015a}, and was first characterized as part of the ALFALFA survey \citep{Giovanelli2013,Rhode2013,Skillman2013,McQuinn2013,Bernstein-Cooper2014}. These studies have characterized its star formation history as low and roughly continuous over the age of the Universe \citep[$<$SFR$>$ = $4.3\times 10^{-5}$~M$_{\odot}$~yr$^{-1}$,][]{McQuinn2015a}, its stellar mass \citep[$M_{*} = 5.7 \times 10^{5}$~M$_{\odot}$,][]{McQuinn2013}, its \hi and dynamical mass \citep[$M_{\rm HI}(r<r_{\rm HI}) = 9.5\times 10^{5}$~M$_{\odot}$, $M_{\rm dyn}(r<r_{\rm HI}) = 2.6 \times 10^{7}$~M$_{\odot}$, $r_{\rm HI} \sim 500$~pc,][]{Bernstein-Cooper2014}, and gas-phase abundances from an \hii region \citep[12 + log(O/H) = 7.17$\pm$0.04][]{Skillman2013}. HI, optical, and H$\alpha$ observations of Leo P are shown inf Figure~\ref{intro:fig:Leo P}.

@@

One of the most common ways to model and understand the stellar populations of galaxies in high-dimensional chemical space is through simplified one-zone (or many-zone) models. This treatment extends back many decades \citep[e.g.][]{Schmidt1963,TalbotArnett1971,Lynden-Bell1975}, where galaxies and their total metal abundances were treated in closed-box systems of gas and stars. Over the intervening decades these models have improved, including prescriptions for gas outflow (a ``leaky-box") or gas inflow (an ``accreting box"), or both \citep[a ``bathtub" model, e.g.][]{FinlatorDave2008,Bouche2010}. The assumptions that go into these models vary dramatically with complexity. Simple models can assume, for example, that long-lived low mass stars do not contribute to metal enrichment while metals from massive stars are instantaneously recycled into the galaxy's gas reservoir (i.e. ignoring stellar lifetimes) or that all metals mix instantly and homogeneously. More complex models account for individual stellar lifetimes (to some degree), or may try and account for some aspects of inhomogeneous mixing by constructing multi-zone models that may separate a galaxy in radial bins (with or without mixing between bins), may account for a hot and cold component separately (with mixing between the two), or account for separate ISM and CGM components. The output of these models is typically the mean metal abundance evolution within each of the tracked components. There is generally no self-consistent prescription to account for the scatter in any abundance relationship in these models, which, when needed, is often added ad-hoc in post-processing. Regardless of the assumptions that go into these models, they are powerful tools that can be used to probe the general physical processes that govern galactic chemical evolution, and explore the vast, uncertain parameter space associated with the models that characterize these processes \citep[e.g.][]{Cote2017a}.
